Continuous developmental changes in word recognition across early childhood support early language learning

### 2025-02-05 paper/paper.Rmd
more plotting tweaks -- growth curves and arrows

@@

graph_sem(model = fit3_age, text_size = 2, layout = t(layout_age))

nodes <- get_nodes(fit3_age)

edges <- get_edges(fit3_age) |> filter(label!="1.00") |>

graph_sem(edges=edges, nodes=nodes, text_size = 2, layout = t(layout_age),
          rect_width=2,
          rect_height=1.2,
          ellipses_width=1.5,
          ellipses_height=1.5,
          variance_diameter=.9,
          spacing_x=2
)
```

@@

```{r longitudinal, fig.env="figure*", height = 3.5, fig.cap = "Structural equation model showing longitudinal couplings between growth parameters. "}
#layout <- read.csv(here("misc/layout_slopes_observed_growth.csv"), header = FALSE)
#graph_sem(model = slopes_long, text_size = 2, layout=t(layout))
graph_sem(model = slopes_long, text_size = 2)

```

>>>>

```{r growth, fig.cap="TODO veronica messing around. Dark blue is fastest RT, dark red is slowest RT -- not sure what the ideal legend is.", fig.env="figure*", fig.height = 2.5, fig.width=4, out.width=".5\\textwidth"}

library(nlme)

  d_sub_prod <- filter(d_sub, !is.na(prod)) |>
  group_by(subject_id) |>
  mutate(log_rt_0 = log_rt[1],
         acc_0 = long_window_accuracy[1]) |> 
  filter(!is.na(log_rt_0))

mod3 <- nlme(model = prod ~ SSlogis(age, Asym = 1, xmid, scale),
             data = d_sub_prod,
             fixed = xmid + scale ~ log_rt_0,
             random = xmid + scale ~ 1,
             groups = ~ dataset_name,
             start = c(xmid = 20, scale = 3, `xmid:log_rt_0` = 1, `scale:log_rt_0` = 3),
             na.action = na.exclude, 
             control = list(maxIter = 1000, tolerance = 10))

# fixed effects (no CIs)
global_preds <- expand_grid(age = seq(min(d_sub_prod$age), 
                                      max(d_sub_prod$age), .1), 
                            log_rt_0 = c(quantile(d_sub_prod$log_rt, 
                                                  c(.1,.25,.5,.75,.9), 
                                                  na.rm=TRUE)))
global_preds$pred <- predict(mod3, newdata = global_preds, level = 0)

#random effects
dataset_preds <- d_sub |>
  filter(!is.na(prod)) |> # filter on predictor
  group_by(dataset_name) |>
  summarise(min_age = min(age),
            max_age = max(age)) |>
  group_by(dataset_name) |>
  mutate(df = map2(min_age, max_age, ~expand_grid(age = seq(.x, .y, .1), log_rt_0 = c(quantile(d_sub_prod$log_rt, 
                                                  c(.1,.25,.5,.75,.9), 
                                                  na.rm=TRUE))))) |>
  select(-min_age, -max_age) |>
  unnest(col = "df") |>
  ungroup()

library(viridis)
dataset_preds$pred <- predict(mod3,
                              newdata = dataset_preds,
                              level = c(0,1))$predict.dataset_name

ggplot(d_sub_prod, aes(x = age, y = prod, col = as.factor(log_rt_0))) + 
  geom_point(alpha = .02, color="black") + 
  geom_line(data = dataset_preds |>   filter(dataset_name %in% datasets_with_age_variation$dataset_name), aes(y = pred, group = interaction(log_rt_0, dataset_name)), lty="dashed", alpha=.5)+
    geom_line(data = global_preds, aes(y = pred, group = log_rt_0, col = as.factor(log_rt_0)), size=1)+
  scale_color_viridis(discrete=T, option="H")+
  theme(legend.position = "none")+
  labs(y="Production", x="Age in months")
# +
#   geom_line(data = dataset_preds, aes(y = pred))
```

@@

node_text <- tibble(name=c("acc", "acc_sd", "accuracy", "comp", "log_age", "log_rt", "log_rt_sd", "prod", "speed", "vocab"), pretty=c("Accuracy", "Accuracy (sd)", "Accuracy", "Comprehension", "Age (log months)", "log RT", "log RT (SD)", "Production", "Speed", "Vocabulary"))

nodes <- get_nodes(fit3_age) |> tibble() |> left_join(node_text) |> mutate(label=ifelse(shape=="oval",pretty, str_c(pretty, "\n", est_sig)))

edges <- get_edges(fit3_age) |> filter(label!="1.00") |> mutate(label_location=.5) |> mutate(label_hjust=.5, label_vjust=.5, label_lineheight=1.5)

edges[9, "label_hjust"]=1
edges[7, "label_hjust"]=1

edges[18, "label_hjust"]=0
edges[16, "label_hjust"]=0

edges[17,"curvature"]=50
edges[17,"label_location"]=.6

source(here("helper/modified_sem_plot.R"))

#View(edges)
prepare_graph(edges=edges, nodes=nodes, text_size = 2.5, layout = t(layout_age),
          rect_width=4.2,
          rect_height=3,
          ellipses_width=3,
          ellipses_height=3,
          variance_height=2,
          variance_width=1.5,
          arrow_angle=15,
          arrow_length=.1,
          var_arrow_angle=15,
          var_arrow_length=.1,
          spacing_y=4,
          spacing_x=4,
          fix_coord=FALSE) |> plot_happy_arrows()
```

@@

```{r longitudinal, fig.env="figure*", height = 3.5, fig.cap = "Structural equation model showing longitudinal couplings between growth parameters. "}
layout <- read.csv(here("misc/layout_slopes_observed_growth.csv"), header = FALSE)

node_text <- tribble(~name, ~ pretty, ~fill, 
                     "acc_t0", "t0 Accuracy", "lightblue",
                     "acc_t1", "t1 Accuracy", "lightblue",
                     "acc_t2", "t2 Accuracy","lightblue",
                     "acc_t3", "t3 Accuracy","lightblue",
                     "acc_t4", "t4 Accuracy","lightblue",
                     "accuracy_intercept", "Accuracy intercept","lightblue",
                     "accuracy_slope", "Accuracy slope","lightblue",
                     "log_rt_t0", "t0 log RT","pink",
                     "log_rt_t1", "t1 log RT","pink",
                     "log_rt_t2", "t2 log RT","pink",
                     "log_rt_t3", "t3 log RT","pink",
                     "log_rt_t4", "t4 log RT","pink",
                     "prod_t0" , "t0 Production", "lightgreen",
                     "prod_t1" , "t1 Production",  "lightgreen",
                     "prod_t2" , "t2 Production" ,  "lightgreen",
                     "prod_t3", "t3 Production",          "lightgreen",
                     "prod_t4", "t4 Production" , "lightgreen",
                     "speed_intercept" , "Speed intercept" ,  "pink",
                     "speed_slope" , "Speed slope","pink",
                     "vocab_intercept", "Vocabulary intercept", "lightgreen",
                     "vocab_slope", "Vocabulary slope",  "lightgreen" ) |> 
  mutate(alpha=.5)

nodes <- get_nodes(slopes_long) |> tibble() |> left_join(node_text) |> mutate(label=pretty)

edges <- get_edges(slopes_long) |> mutate(label_location=.5) |> mutate(label=ifelse(arrow=="last", NA, label)) |>  mutate(connect_from=ifelse(to==from, "right",NA),
                    connect_to=connect_from) |> 
  mutate(color="black", label_hjust=.5, label_location=.5)

edges[52,"curvature"]=85
edges[52,"color"]="blue"
edges[61,"curvature"]=85
edges[61,"color"]="blue"
edges[66,"curvature"]=85
edges[66,"color"]="blue"
edges[52, "label_hjust"]=0
edges[61, "label_hjust"]=0
edges[66, "label_hjust"]=0

edges[54,"curvature"]=85
edges[54, "label_hjust"]=0

edges[63,"curvature"]=85
edges[63, "label_hjust"]=0

edges[53,"curvature"]=80
edges[53,"color"]="red"
edges[55,"curvature"]=82
edges[55,"label_location"]=.55

edges[55,"color"]="red"
edges[62,"curvature"]=80
edges[62,"color"]="red"
edges[53, "label_location"]=.6
edges[53, "label_hjust"]=0.4

edges[58,"curvature"]=80
edges[58,"color"]="darkgreen"
edges[60,"curvature"]=82
edges[60,"color"]="darkgreen"
edges[60,"label_location"]=.45
edges[65,"curvature"]=80
edges[65,"color"]="darkgreen"
edges[65, "label_location"]=.4
edges[53, "label_hjust"]=0.4

edges[59,"curvature"]=77

edges[64,"curvature"]=75
edges[56,"curvature"]=75
edges[57,"curvature"]=75

source(here("helper/modified_sem_plot.R"))
#View(edges)
prepare_graph(edges=edges, nodes=nodes, text_size = 2, layout = t(layout),angle=0,
          rect_width=1.5,
          rect_height=2,
          ellipses_width=1.5,
          ellipses_height=3,
          variance_width= .5,
          variance_height=1.3,
          var_arrow_angle=10,
          var_arrow_length=.05,
          arrow_angle=15,
          arrow_length=.07,
          variance_diameter=1,
          spacing_y=2.5,
          spacing_x=1.5,
          fix_coord=FALSE) |> 
  edit_edges({connect_from=case_when(
    (to==from & from %in% c("accuracy_intercept","accuracy_slope","speed_slope",
                          "speed_intercept", "vocab_slope", "vocab_intercept"))~"bottom",
    to==from ~ "right",
    !is.na(curvature)~"left",
    T ~ "right")}) |> 
  plot_happy_arrows()
```
